Objectification in Action: Self- and Other-Objectification in Same-gender Interactions

### 2020-07-21 Final_Manuscript.Rmd
fixed bib

Self-objectification has been found to occur after a mere relationship prime among women [@sanchez2008romance] because in our culture women need to look attractive to obtain and maintain successfuol relationship, thus, the "college relationships condition may highten self-objectification and the evaluation of other women in a sexualized way. Researchers have also studied interpersonal objectification and self-objectification in romantic relationships [@strelan2018birds]. Evidence is mixed, and women can sometimes have positive outcomes from appearance valuation within the context of an established romantic relationship [@meltzer2020women], less is known about first dates. But this literature is also only about heterosexual relationships. A recent study has investigated the effect of self-objectification on the reduction of relationship biulding skills in general (including same-sex friendships) [CITE PWQ paper I reviewed, Yoder editor]. Empirical evidence reveals that objectification manifests through inauthenticity in romantic relationships [@brunelletal2010], adverse attitudes in regard to career aspirations, and a decrease in concentration and impairment in female cognitive performance [@quinnetal].

>>>>

Self-objectification has been found to occur after a mere relationship prime among women [@sanchez2008romance] because in our culture women need to look attractive to obtain and maintain successfuol relationship, thus, the "college relationships condition may highten self-objectification and the evaluation of other women in a sexualized way. Researchers have also studied interpersonal objectification and self-objectification in romantic relationships [@strelan2018birds]. Evidence is mixed, and women can sometimes have positive outcomes from appearance valuation within the context of an established romantic relationship and they experience less SSO after appearance comments within romantic relationships [@meltzer2020women], less is known about first dates. But this literature is also only about heterosexual relationships. A recent study has investigated the effect of self-objectification on the reduction of relationship biulding skills in general (including same-sex friendships) [CITE PWQ paper I reviewed, Yoder editor]. Empirical evidence reveals that objectification manifests through inauthenticity in romantic relationships [@brunelletal2010], adverse attitudes in regard to career aspirations, and a decrease in concentration and impairment in female cognitive performance [@quinnetal].

### 2020-07-21 Final_Manuscript.tex
fixed bib

Theorists across diverse disciplines have explored the multiple ways
that the body conveys social meaning, as well as how these meanings
shape gendered experiences, especially during interpersonal interactions
with members of different social groups (``Engendered lives,'' 1993,
Garcia, Earnshaw, and Quinn (2016)). Objectification theory, a valuable
theoretical framework proposed by Roberts and Fredrickson (1997) which
works to contextualize female bodies as socio-cultural constructions,
can be employed to illuminate the gender oppression and negative lived
experiences of girls and women. Empirical evidence illustrates how women
continue to be objects of interpersonal discrimination and expereince
daily sexist hassles (Swim, Hyers, Cohen, \& Ferguson, 2001). One form
of interpersonal discrimination women face is the process by which their
whole being is viewed as a collection of sexualized body parts valued
predominantly for commodification, a phenomena termed sexual
objectification (Bartky, 1990). Sexual objectification occurs with both
\enquote{endless variety and monotonous similarity,} and is thus
mediated by unique combinations of race, ethnicity, sexuality, age, and
class (Fredrickson, Hendler, Nilsen, O'Barr, \& Roberts, 2011; Rubin,
1975, cited in Fraser and Nicholson (1989), p.~28). Amid such
heterogeneity though, \enquote{having a reproductively mature female
body} proposed by Roberts and Fredrickson (1997) is likely to create a
shared vulnerability to sexual objectification and a variety of shared
negative experiences as a result.

Self-objectification is a multidimensional process that accounts for the
cognitive mechanism that translates experiences of sexualization at the
cultural level to psychological (e.g., anxiety, self-esteem,
authenticity, motivational states) and behavioral (cognitive
performance, body monitoring) features of mental health and well-being
at the individual level (Calogero, Tantleff-Dunn, \& Thompson, 2011;
Moradi \& Huang, 2008). Calogero et al. (2011) proposes that the
construct of self-objectification can be conceptualized as a learned
trait. Furthermore, it can also be elicited momentarily, through the
media, for example, with sexualized images in movies and magazines,
which can lead to a state of self-objectification (Calogero et al.,
2011, Moradi and Huang (2008)). Being objectified by another person and
possessing trait-level self-objectification (TSO) may interact to
influence experiences of feeling like a body, or state
self-objectification (SSO) (CITE swimsuit sweater, Garcia et al., 2016).

Studies have shown that within social encounters women are gazed at more
than men (Briton \& Hall, 1995, CITE saguy skype study, calogero video
walking study), often times feel \enquote{looked at} within
interpersonal interactions (Argyle \& Williams, 1969), and will more
than likely internalize the objectifying gaze on physical self (Young,
1979). Moreover, perhaps the most adverse effect of objectifying
treatment is that it effectively socializes girls and women to treat
themselves as objects to be looked at and evaluated, an effect termed
self-objectification (Bartky, 1990; Berger, Cohen, \& Zelditch Jr, 1972;
L. Fredrickson, Roberts, M. Noll, Quinn, \& Twenge, 1998).

ENTIRE PARAGRAPH ON MODERN (POST-2016) STUDIES OF INTERPERSONAL
OBJECTIFICATION.

Empirical evidence reveals that objectification manifests through
inauthenticity in romantic relationships (Brunell et al., 2010), adverse
attitudes in regard to career aspirations, and a decrease in
concentration and impairment in female cognitive performance (D. M.
Quinn, Chaudoir, \& Kallen, 2011).

@@

@@

As stated previously, we did not find a significant effect between actor
SSO and actor authenticity (\(\beta\)= -.10, \(p\)=.38), which suggests
that there is not sufficient evidence to support the claim that partner
objectification is the cause for the diverse range of negative effects
related to relationship inauthenticity. However, we did observe a
significant partner effect of objectification on actor
self-objectification, which does align with our hypothesis that
theorized objectification from other women can also cause women to
self-objectify just as they do within interactions between male-female
pairs. The results of this study demonstrate the complex and ambivalent
nature of female sexual objectification and additionally highlight the
psychological and social consequences of such objectification processes
on women's social relationships and well-being. The sample of the
current study was comprised of Western women, being that sexual
objectification is most prevalent in this culture (Loughnan et al.,
2015), and research on objectification conducted outside of Western or
Westernized countries is scarce (Moradi \& Huang, 2008). Because
\enquote{bodies exist within social and cultural contexts, and hence are
also constructed through sociocultural practices and discourses}
(Roberts \& Fredrickson, 1997, p. 174), it is important to consider how
diverse social identities within unique cultural contexts may inform
sexual objectification phenomenon to test the cross-cultural
applicability of theoretical frameworks (Loughnan et al., 2015).
Further, sexualzing experiences and self-objectification are thought to
begin a very young age, and thus, researchers have only recently begun
to examine such experiences among children (Bury, Tiggemann, \& Slater,
2016; e.g., Holland \& Haslam, 2016; Jongenelis, Byrne, \& Pettigrew,
2014). Considering the fact that the average mean age of the
investigated participants of this current study was 18.85 years,
research among younger and older individuals is needed, especially
because self-objectification may change over time (Roberts \&
Fredrickson, 1997). It may be valuable to question the extent to which
children, adolescents, or emerging adults of different races or
ethnicities are exposed to varied amounts of sexulizing content. Also,
future experiments or longitudinal studies should explore the external
validity of the notions of self-objectification and how the
operalization of self-objectification may be improved. Regardless, the
results from the current analysis highlight how subtle forms of sexist
discrimination operate to inform prevention and intervention efforts in
both clinical and educational contexts. These results are quite useful
for promoting mental health and within early action programs for girls
and young women, where scholars and practitioners might provide the
tools necessary to circumvent or mitigate negative effects on
self-objectification, and combat such experiences.

@@

\hypertarget{ref-Kaschak}{}
Engendered lives: A new psychology of womens experience. (1993).
\emph{Choice Reviews Online}, \emph{30}(07).
doi:\href{https://doi.org/10.5860/choice.30-4102}{10.5860/choice.30-4102}

@@

@@

\hypertarget{ref-loughnan2015}{}
Loughnan, S., Fernandez-Campos, S., Vaes, J., Anjum, G., Aziz, M.,
Harada, C., \ldots{} Tsuchiya, K. (2015). Exploring the role of culture
in sexual objectification: A seven nations study. \emph{Revue
Internationale de Psychologie Sociale}, \emph{28}(1), 125--152.

@@

@@

@@

@@

>>>>

OUTLINE - Introduction: Objectification theory says stuff. Recently
we've learned about interpersonal objectification but we only know it
about mixed gender. We know women objectify each other, but what about
in interpersonal interactions? - Self-objectification - SSO versus TSO -
Does this process of self-objectification also lead women to objectify
other women? - Women objectifying other women - Lots of evidence, but we
don't know if this happens interpersonally - Interpersonal
objectifcation research is a whole thing now, but there are no studies
(or theoretical models) that include women-on-women interpersonal
objectification processes. - Relationships and objectification - Both
romantic relationships and relationships in general have been studies in
the context of objectification, but neither of these literatures have
investigated specifically same-gender relationships. - Authenticity
paragraph? - Current Study - Alternative predictions

Roberts and Fredrickson (1997)\enquote{s Objectification Theory,
suggests that women are objectified in interpersonal encounters, and
that the effect might be strongest from men, but possibly also from
women. Psychological researchers studying the sexual objectification of
women, and the resulting self-objectification that follows, have in
recent years explored the interpersonal process of the objectification -
self-objectification link. For exmaple, in a face-to-face mixed-gender
dyadic study, Garcia, Earnshaw, and Quinn (2016) found that men's
reported objectificaiton of their female interaction partner during an
interaction was associated with increased sefl-objectification reported
by their partner. Gervais, Sáez, Riemer, and Klein (2020) has reviewed
the literature on interpersonal self-objectification and has organized
the the process in a theoretical model called the SIMO, but this model
is only relevant for mixed-gender encouters. Although there is ample
evidence that women objectify other women (Loughnan et al., 2015a; Puvia
\& Vaes, 2013), there is currently no studies investigating the process
of interpersonal objectification same-gender interactions. Does
objectification by a female interaction partner have the same downstream
negative consequences as being objectified by a male interaction
partner? For example, perhaps women could be amplifying each others}
self-objectification as they get ready to go out for the night,
\emph{before} the theoretical encounters with men outlined in Gervais et
al. (2020) occur.

Self-objectification is a multidimensional process that accounts for the
cognitive mechanism that translates experiences of sexualization at the
cultural level to psychological (e.g., anxiety, self-esteem,
authenticity, motivational states) and behavioral (cognitive
performance, body monitoring) features of mental health and well-being
at the individual level (Calogero, Tantleff-Dunn, \& Thompson, 2011;
Moradi \& Huang, 2008). Calogero et al. (2011) proposes that the
construct of self-objectification can be conceptualized as a learned
trait. Furthermore, it can also be elicited momentarily, through the
media, for example, with sexualized images in movies and magazines,
which can lead to a state of self-objectification (Calogero et al.,
2011, Moradi and Huang (2008)). Being objectified by another person and
possessing trait-level self-objectification (TSO) may interact to
influence experiences of feeling like a body, or state
self-objectification (SSO) (CITE swimsuit sweater, Garcia et al., 2016).
Does this habit self-objectification lead women to also objectify other
women, possibly to compare oneself?

Some studies have found that women objectify other women to a greater
extent than they objectify men (Loughnan et al., 2015a). Women objectify
(dehumanized) sexualized targets presented as images (Puvia \& Vaes,
2013), but we know next to nothing about interpersonal sexual
objectification among women and what the immediate effects might be.
Further, do women objectify non-sexualized women they are interacting
with? What might be different for same-gender interactions? Harsey and
Zurbriggen (2020) found that self-objectification was related to the
objectification of women to a similar degree for men and women
participants, although men scored high than women on a measure of belief
that objectification of women is natural for men.

SIMO model and literature review (Gervais et al., 2020). Studies have
shown that within social encounters women are gazed at more than men
(Briton \& Hall, 1995, CITE saguy skype study, calogero video walking
study), often times feel \enquote{looked at} within interpersonal
interactions (Argyle \& Williams, 1969), and will more than likely
internalize the objectifying gaze on physical self (Young, 1979).
Moreover, perhaps the most adverse effect of objectifying treatment is
that it effectively socializes girls and women to treat themselves as
objects to be looked at and evaluated, an effect termed
self-objectification (Bartky, 1990; Berger, Cohen, \& Zelditch Jr, 1972;
L. Fredrickson, Roberts, M. Noll, Quinn, \& Twenge, 1998). We don't know
about women-on-women interpersonal objectification. What might be
different?

Self-objectification has been found to occur after a mere relationship
prime among women (Sanchez \& Broccoli, 2008) because in our culture
women need to look attractive to obtain and maintain successfuol
relationship, thus, the ``college relationships condition may highten
self-objectification and the evaluation of other women in a sexualized
way. Researchers have also studied interpersonal objectification and
self-objectification in romantic relationships (Strelan \& Pagoudis,
2018). Evidence is mixed, and women can sometimes have positive outcomes
from appearance valuation within the context of an established romantic
relationship and they experience less SSO after appearance comments
within romantic relationships (Meltzer, 2020), less is known about first
dates. But this literature is also only about heterosexual
relationships. A recent study has investigated the effect of
self-objectification on the reduction of relationship biulding skills in
general (including same-sex friendships) {[}CITE PWQ paper I reviewed,
Yoder editor{]}. Empirical evidence reveals that objectification
manifests through inauthenticity in romantic relationships (Brunell et
al., 2010), adverse attitudes in regard to career aspirations, and a
decrease in concentration and impairment in female cognitive performance
(D. M. Quinn, Chaudoir, \& Kallen, 2011).

MAYBE SEPARATE PARAGRAPH ABOUT AUTHENTICITY LINK SPECIFICALLY. And how
woman-woman interpersonal objectificaiton preocesses might diverge from
this process. {[}again CITE PWQ paper I reviewed, Yoder editor{]}.

@@

Puvia and Vaes (2013) would alternatively predict that women's tendency
to self-objectify (TSO) leads them to objectify other women
(\ldots{}well, dehumanize a sexualized woman, but also Strelan and
Hargreaves (2005) would predict the TSO-\textgreater{}SOO link and
Harsey and Zurbriggen (2020) too), and this is mediated by the woman's
own state self-objectification (SSO). So we also tested this as a
possible alternative model.

@@

As stated previously, we did not find a significant effect between actor
SSO and actor authenticity, which suggests that there is not sufficient
evidence to support the claim that partner objectification is the cause
for the diverse range of negative effects related to relationship
inauthenticity. However, we did observe a significant partner effect of
objectification on actor self-objectification, which does align with our
hypothesis that theorized objectification from other women can also
cause women to self-objectify just as they do within interactions
between male-female pairs.

\subsection{Sexism and Sexual
Discrimination}\label{sexism-and-sexual-discrimination}

Sexism has recently been found to be related to the objectification of
women, but not men, for both men and women (Harsey \& Zurbriggen, 2020).
Empirical evidence illustrates how women continue to be objects of
interpersonal discrimination and expereince daily sexist hassles (Swim,
Hyers, Cohen, \& Ferguson, 2001). One form of interpersonal
discrimination women face is the process by which their whole being is
viewed as a collection of sexualized body parts valued predominantly for
commodification, a phenomena termed sexual objectification (Bartky,
1990). Sexual objectification occurs with both \enquote{endless variety
and monotonous similarity,} and is thus mediated by unique combinations
of race, ethnicity, sexuality, age, and class (Fredrickson, Hendler,
Nilsen, O'Barr, \& Roberts, 2011; Rubin, 1975, cited in Fraser and
Nicholson (1989), p.~28). Amid such heterogeneity though,
\enquote{having a reproductively mature female body} proposed by Roberts
and Fredrickson (1997) is likely to create a shared vulnerability to
sexual objectification and a variety of shared negative experiences as a
result.

\subsection{Limitations and Future
Directions}\label{limitations-and-future-directions}

\subsubsection{Sample Characteristics}\label{sample-characteristics}

The sample of the current study was comprised of Western women, being
that sexual objectification is most prevalent in this culture (Loughnan
et al., 2015b), and research on objectification conducted outside of
Western or Westernized countries is scarce (Moradi \& Huang, 2008).
Because \enquote{bodies exist within social and cultural contexts, and
hence are also constructed through sociocultural practices and
discourses} (Roberts \& Fredrickson, 1997, p. 174), it is important to
consider how diverse social identities within unique cultural contexts
may inform sexual objectification phenomenon to test the cross-cultural
applicability of theoretical frameworks (Loughnan et al., 2015b).
Further, sexualzing experiences and self-objectification are thought to
begin a very young age, and thus, researchers have only recently begun
to examine such experiences among children (Bury, Tiggemann, \& Slater,
2016; e.g., Holland \& Haslam, 2016; Jongenelis, Byrne, \& Pettigrew,
2014). Considering the fact that the average mean age of the
investigated participants of this current study was 18.85 years,
research among younger and older individuals is needed, especially
because self-objectification may change over time (Roberts \&
Fredrickson, 1997). It may be valuable to question the extent to which
children, adolescents, or emerging adults of different races or
ethnicities are exposed to varied amounts of sexulizing content.

\subsubsection{Measurement}\label{measurement}

Also, future experiments or longitudinal studies should explore the
external validity of the notions of self-objectification and how the
operalization of self-objectification may be improved.

\subsubsection{Alternative Model}\label{alternative-model}

Puvia and Vaes (2013) alternative model. TSO -\textgreater{} SOO,
mediated by SSO.

\subsection{Clinical Implications}\label{clinical-implications}

Regardless, the results from the current analysis highlight how subtle
forms of sexist discrimination operate to inform prevention and
intervention efforts in both clinical and educational contexts. These
results are quite useful for promoting mental health and within early
action programs for girls and young women, where scholars and
practitioners might provide the tools necessary to circumvent or
mitigate negative effects on self-objectification, and combat such
experiences.

\section{Conclusion}\label{conclusion}

The results of this study demonstrate the complex and ambivalent nature
of female sexual objectification and additionally highlight the
psychological and social consequences of such objectification processes
on women's social relationships and well-being.

@@

@@

\hypertarget{ref-gervais2020social}{}
Gervais, S. J., Sáez, G., Riemer, A. R., \& Klein, O. (2020). The social
interaction model of objectification: A process model of goal-based
objectifying exchanges between men and women. \emph{British Journal of
Social Psychology}, \emph{59}(1), 248--283.

\hypertarget{ref-harsey2020men}{}
Harsey, S. J., \& Zurbriggen, E. L. (2020). Men and women's
self-objectification, objectification of women, and sexist beliefs.
\emph{Self and Identity}, 1--8.

@@

\hypertarget{ref-loughnan2015exploring}{}
Loughnan, S., Fernandez-Campos, S., Vaes, J., Anjum, G., Aziz, M.,
Harada, C., \ldots{} Tsuchiya, K. (2015a). Exploring the role of culture
in sexual objectification: A seven nations study. \emph{Revue
Internationale de Psychologie Sociale}, \emph{28}(1), 125--152.

\hypertarget{ref-loughnan2015}{}
Loughnan, S., Fernandez-Campos, S., Vaes, J., Anjum, G., Aziz, M.,
Harada, C., \ldots{} Tsuchiya, K. (2015b). Exploring the role of culture
in sexual objectification: A seven nations study. \emph{Revue
Internationale de Psychologie Sociale}, \emph{28}(1), 125--152.

@@

\hypertarget{ref-meltzer2020women}{}
Meltzer, A. L. (2020). Women can benefit from sexual and physical
valuation in the context of a romantic relationship. \emph{Personality
and Social Psychology Bulletin}, \emph{46}(2), 243--257.

@@

\hypertarget{ref-puvia2013being}{}
Puvia, E., \& Vaes, J. (2013). Being a body: Women's appearance related
self-views and their dehumanization of sexually objectified female
targets. \emph{Sex Roles}, \emph{68}(7-8), 484--495.

@@

\hypertarget{ref-sanchez2008romance}{}
Sanchez, D. T., \& Broccoli, T. L. (2008). The romance of
self-objectification: Does priming romantic relationships induce states
of self-objectification among women? \emph{Sex Roles}, \emph{59}(7-8),
545--554.

@@

\hypertarget{ref-strelan2005women}{}
Strelan, P., \& Hargreaves, D. (2005). Women who objectify other women:
The vicious circle of objectification? \emph{Sex Roles},
\emph{52}(9-10), 707--712.

\hypertarget{ref-strelan2018birds}{}
Strelan, P., \& Pagoudis, S. (2018). Birds of a feather flock together:
The interpersonal process of objectification within intimate
heterosexual relationships. \emph{Sex Roles}, \emph{79}(1-2), 72--82.
